Reproducible Research: Peer Assessment 1
========================================
```{r dataTime,echo=FALSE}
time <- format(Sys.time(), "%a %dth of %b in %Y")
```

 by Juan M. Medel on `r time`
-----------------------------
 
## 1. Loading and preprocessing the data
I'll load the data and remove the element which has no valid data in the steps column
```{r LoadingData}
rawData <- read.csv(file="activity.csv") 
head(PreProcesedData)
```
## 2. What is mean total number of steps taken per day?
```{r barplot2, fig.width=20}
stepsPerDay <- tapply(X = rawData$steps, INDEX=rawData$date, FUN = sum)
par(las = "2", mai = c(1.660000, 1.093333, 1.093333, 0.560000))
barplot(stepsPerDay,width=5 ,main="steps per day", xlab ="day", ylab = "Total steps")
stepmean <- mean(stepsPerDay,na.rm=T)
stepmedian <- median(stepsPerDay, na.rm=T)
c(stepmean,stepmedian)
```
The mean of the steps taken per day is `r stepmean` and the median is `r stepmedian`


## 3. What is the average daily activity pattern?
```{r barplot3, fig.width=20}
stepsPerPeriod <- tapply(X = rawData$steps, INDEX=rawData$interval, FUN = mean, na.rm=TRUE)
library("ggplot2")
qplot(x=as.numeric(names(stepsPerPeriod)),y=stepsPerPeriod, geom="line") + 
    labs(title = "step average per period", x = "5 minutes period", y = "Mean steps")
MaxSteps <- round(stepsPerPeriod[order(stepsPerPeriod, decreasing=T)[1]])
MaxStepPeriod <- names(MaxSteps)
#convert the character in a valid hours
time <- paste( as.character(trunc(as.numeric(MaxStepPeriod)/100)),
               as.character(as.numeric(MaxStepPeriod) - 100*trunc(as.numeric(MaxStepPeriod)/100)), 
               sep = ":")
```
The maximun average steps is `r MaxSteps` that happen at `r time`  


## 4. Imputing missing values
For every missing value, the inputed value will be the average steps of this period 
```{r missingValues}
completedData <- rawData
# if the data is missing then fullfilled with the mean for the period of days with data calculated before
for(i in 1:dim(completedData)[1]){
    if(is.na(completedData[i,1])){
        completedData[i,1] <- stepsPerPeriod[as.character(completedData[i,3])]
    }
}
totalMissingValuesRaw <- sum(is.na(rawData$steps))
totalMissingValuesCompleted <- sum(is.na(completedData$steps))
```
The total NA values are `r totalMissingValuesRaw` and finally the missing values on the processed data is `r totalMissingValuesCompleted`

```{r Comparation, fig.width=20}
stepsPerDayCompleted <- tapply(X = completedData$steps, INDEX=completedData$date, FUN = sum)
par(las = "2", mai = c(1.660000, 1.093333, 1.093333, 0.560000))
barplot(stepsPerDayCompleted,width=5 ,main="steps per day (NA's filled)", 
        xlab ="day", ylab = "Total steps")
stepmeanCompleted <- mean(stepsPerDayCompleted)
stepmedianCompleted <- median(stepsPerDayCompleted)
c(stepmeanCompleted,stepmedianCompleted)
totalStepPerDay <- sum(stepsPerDay, na.rm=T)
totalStepPerDayCompleted <- sum(stepsPerDayCompleted)
c(totalStepPerDay, totalStepPerDayCompleted)
```
The mean steps per day in raw data is `r stepmean`and the steps mean per day in completed data is `r stepmeanCompleted`. As we can see the mean keep been the same because we remove the NA's previously while calculating the mean.

The median steps per day in wad data is `r stepmedian` and the steps meadian per day in completed data is `r stepmedianCompleted`. As we can se the value practicaly the same value.

There is no impact in the data because the filling values are made with the averange of steps a normal day. Nevertheless the total steps per days in raw data is `r totalStepPerDay` and in the fullfilled data is `r totalStepPerDayCompleted` that is higher because of the adding values that previusly are missed.

## 5. Are there differences in activity patterns between weekdays and weekends?
```{r weekdaysWeekendsComparation, fig.width=20}
#Create a new dataframe with two more columns, one for the weekdayname and other for the factor
rawDataWeekday <- cbind(rawData, weekdays(as.Date(rawData[,2])), numeric(dim(rawData)[1]))
names(rawDataWeekday)[4] <- "weekday"
names(rawDataWeekday)[5] <- "factorweekday"
#Create a vector with the names of the weekdays (valid for every languaje)
daysnames <- weekdays(as.Date(levels(rawDataWeekday[,2])[1:7]))
weekdays <- daysnames[1:5]
weekend <- daysnames[6:7]
#assign the factor weekday or weekend for the current day in column 2
for(i in 1:dim(rawDataWeekday)[1]){
    if(rawDataWeekday[i,4] %in% weekdays) rawDataWeekday[i,5] <- 1
    if(rawDataWeekday[i,4] %in% weekend) rawDataWeekday[i,5] <- 2
}
rawDataWeekday$factorweekday <- as.factor(x=rawDataWeekday$factorweekday)
levels(rawDataWeekday$factorweekday) <- c("weekdays", "weekend")
#split the data on weekend and weekdays for the interested data (steps and interval) and calculate the mean of the steps in this perior along the days
weekdaysStepPeriod <- rawDataWeekday[rawDataWeekday$factorweekday=="weekdays",c(1,3)]
stepsPerPeriodWeekday <- tapply(X = weekdaysStepPeriod$steps, INDEX=weekdaysStepPeriod$interval,
                                FUN = mean, na.rm=TRUE)
weekendStepPeriod <- rawDataWeekday[rawDataWeekday$factorweekday=="weekend",c(1,3)]
stepsPerPeriodWeekend <- tapply(X = weekendStepPeriod$steps, INDEX=weekendStepPeriod$interval,
                                FUN = mean, na.rm=TRUE)
#Create a dataframe that hold the information in ordet to create the plot
stepsPerPeriodWeekday <- cbind(names(stepsPerPeriodWeekday),stepsPerPeriodWeekday,character(length(stepsPerPeriodWeekday)))
stepsPerPeriodWeekday <- as.data.frame(stepsPerPeriodWeekday)
names(stepsPerPeriodWeekday) <- c("period", "steps", "factorWeek")
stepsPerPeriodWeekday$factorWeek <- "weekday"

stepsPerPeriodWeekend <- cbind(names(stepsPerPeriodWeekend),stepsPerPeriodWeekend,character(length(stepsPerPeriodWeekend)))
stepsPerPeriodWeekend <- as.data.frame(stepsPerPeriodWeekend)
names(stepsPerPeriodWeekend) <- c("period", "steps", "factorWeek")
stepsPerPeriodWeekend$factorWeek <- "weekend"
#joint the complete dataframe:
stepsPerPeriodFatorWeek <- rbind(stepsPerPeriodWeekday,stepsPerPeriodWeekend)
stepsPerPeriodFatorWeek$factorWeek <- as.factor(stepsPerPeriodFatorWeek$factorWeek)
stepsPerPeriodFatorWeek$period <- as.numeric(stepsPerPeriodFatorWeek$period)
stepsPerPeriodFatorWeek$steps <- as.numeric(stepsPerPeriodFatorWeek$steps)
qplot(x = period, y = steps, data = stepsPerPeriodFatorWeek, facets = factorWeek ~ .,) + 
    labs(title = "step average per period", x = "5 minutes period", y = "Mean steps") + 
    geom_smooth(method = "auto", se=FALSE, col="steelblue") + geom_line()
stepsMeanWeekday <- mean(as.numeric(stepsPerPeriodWeekday$steps), na.rm=T)
stepsMeanWeekend <- mean(as.numeric(stepsPerPeriodWeekend$steps), na.rm=T)
c(stepsMeanWeekday, stepsMeanWeekend)
```
As you can see on the resume the total steps in weekdays is `r stepsMeanWeekday` that is bigger than in weekend with `r stepsMeanWeekend` and the graphic show that the deviation is higher in weekend that in weekdays