Reproducible Research: Peer Assessment 1
========================================
```{r dataTime,echo=FALSE}
time <- format(Sys.time(), "%a %dth of %b in %Y")
```

 by Juan M. Medel on `r time`
-----------------------------
 
## 1. Loading and preprocessing the data
I'll load the data and remove the element which has no valid data in the steps column
```{r LoadingData}
rawData <- read.csv(file="activity.csv") 
PreProcesedData <- rawData[!is.na(rawData$steps),]
head(PreProcesedData)
```
## 2. What is mean total number of steps taken per day?
```{r barplot, fig.width=20}
stepsPerDay <- tapply(X = PreProcesedData$steps, INDEX=PreProcesedData$date, FUN = sum)
par(las = "2", mai = c(1.660000, 1.093333, 1.093333, 0.560000))
barplot(stepsPerDay,width=5 ,main="steps per day", xlab ="day", ylab = "Total steps")
stepmean <- mean(stepsPerDay,na.rm=T)
stepmedian <- median(stepsPerDay, na.rm=T)
c(stepmean,stepmedian)
```
The mean of the steps taken per day is `r stepmean` and the median is `r stepmedian`


## 3. What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
